Constructing Biological Regulatory Networks from Process Hitting models \\\rewriteil{Or possibly: “Complementarity between PH, IG and Thomas modeling”}

### 2013-07-04 parts/infer-ig.tex
finally, switch back and shrink

\pref{pps:inference-edges} details the inference of all existing influences between components occurring
with a threshold $t$.
The main idea of this inference is that
the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
(\pref{eq:edges-inference}).
Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
Furthermore, studying the influences of the groups of regulators of $a$
allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
We ignore the cases where a component has no visible influence on another.
%
\begin{proposition}[Influences inference]\label{pps:inference-edges}
We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
\begin{align}
\begin{split}\label{eq:edges-inference}
  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}\\
\begin{split}\label{eq:edges-inference-auto}
  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}\\
\begin{split}\label{eq:edges-inference-noreg}
  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}
%\begin{split}\label{eq:edges-inference-pos}
%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
%  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j < a_k
%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
%%    &\qquad \warnul{a_j < a_k \quad^{***}}
%\end{split}\\
%\begin{split}\label{eq:edges-inference-neg}
%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
%  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j > a_k
%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
%    %&\qquad \warnul{a_j > a_k \quad^{***}}
%\end{split}
\end{align}
where $\f{sign}(x) = \begin{cases}\frac{|x|}{x} & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
\end{proposition}
%
%\todo{\underline{$^*$Strict inference}}\\
%\todo{\underline{$^{**}$Large inference}}\\
%\todo{\underline{$^{***}$Non-unitary-style inference}}

>>>>

\pref{pps:inference-edges} details the inference of all existing influences between components occurring
with a threshold $t$.
The main idea of this inference is that
the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
(\pref{eq:edges-inference}).
Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
%Furthermore, studying the influences of the groups of regulators of $a$
%allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
%Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
We ignore the cases where a component has no visible influence on another.
%
\begin{proposition}[Influences inference]\label{pps:inference-edges}
We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
\begin{align}
\begin{split}\label{eq:edges-inference}
  \forall b\in\PHpredecgene{a} \cup \{ a \}, \forall s \in \{ +, - \}, \\
  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}\\
%
%%% Tentative de réduire le nombre d'auto-arcs superflus
%
%\begin{split}\label{eq:edges-inference-auto}
%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
%  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
%    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
%    &\qquad s = \f{sign}(a_k - a_j)
%\end{split}\\
%\begin{split}\label{eq:edges-inference-noreg}
%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
%    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
%    &\qquad s = \f{sign}(a_k - a_j)
%\end{split}
%
%%% Ancienne version de l'inférence
%
%\begin{split}\label{eq:edges-inference-pos}
%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
%  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j < a_k
%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
%%    &\qquad \warnul{a_j < a_k \quad^{***}}
%\end{split}\\
%\begin{split}\label{eq:edges-inference-neg}
%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
%  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j > a_k
%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
%    %&\qquad \warnul{a_j > a_k \quad^{***}}
%\end{split}
\end{align}
where $\f{sign}(x) = \begin{cases}\frac{|x|}{x} & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
\end{proposition}
%
%\todo{\underline{$^*$Strict inference}}\\
%\todo{\underline{$^{**}$Large inference}}\\
%\todo{\underline{$^{***}$Non-unitary-style inference}}
